Estimating Fixed-frame Galaxy Magnitudes in the SDSS Spectroscopic Survey$^1$

### 2002-04-14 docs/paper/paper.tex
changed
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@92 75dff457-8751-0410-ab1a-cd701654ce6b

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{#1}{#2}}}

@@

>>>>

\newcommand{\vv}[1]{{\bf #1}}
\newcommand{\df}{\delta}
\newcommand{\dfft}{{\tilde{\delta}}}
\newcommand{\betaft}{{\tilde{\beta}}}
\newcommand{\erf}{{\mathrm{erf}}}
\newcommand{\erfc}{{\mathrm{erfc}}}
\newcommand{\Step}{{\mathrm{Step}}}
\newcommand{\ee}[1]{\times 10^{#1}}
\newcommand{\avg}[1]{{\langle{#1}\rangle}}
\newcommand{\Avg}[1]{{\left\langle{#1}\right\rangle}}
\def\simless{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'074$}}\mathchar"7218\,$}}} % < or of order
\def\simgreat{\mathbin{\lower 3pt\hbox
	{$\,\rlap{\raise 5pt\hbox{$\char'076$}}\mathchar"7218\,$}}} % > or of order
\newcommand{\iras}{{\sl IRAS\/}}
\newcommand{\petroratio}{{{\mathcal{R}}_P}}
\newcommand{\petroradius}{{{r}_P}}
\newcommand{\petronumber}{{{N}_P}}
\newcommand{\petroratiolim}{{{\mathcal{R}}_{P,\mathrm{lim}}}}
\newcommand{\band}[2]{\ensuremath{^{#1}\!{#2}}}

@@

In Figure \ref{response_sdss}, there is clearly a gap between the $g$
and $r$ bands. Linear fits to the data allow considerable freedom in
the resulting reconstructed SEDs. As it happens, one of the directions
in our best fit coefficient space $a_i/a_0$ corresponds to a large
spike at around 4000 \AA; that this direction is important is not
surprising, since one of the most variable qualities of galaxy SEDs is
the size of the 4000 \AA\ break. However, the gap between the $g$ and
$r$ bands corresponds to the 4000 \AA\ break at around $z=0.3$. The
existence of this spike in one direction in our four-dimensional space
means that galaxies at $z=0.3$ can vary along this direction in order
to better fit $u$, $i$, and $z$, with little effect on the $g-r$ color
of the fit. This results in unphysical SED fits to galaxies near
$z=0.3$.

To demonstrate this degeneracy, Figure \ref{spur} shows the spectrum
corresponding to the direction represented by the vector $-0.6
\vv{e}_1(\lambda) + \vv{e}_3$. We overplot the \band{0.3}{g} and
\band{0.3}{r} bands. This direction has a strong peak in the gap
between the two bands at $z=0.3$.  This difficulty is why, in Section
\ref{data}, we exclude the regime around $z\sim 0.3$; otherwise the
templates became distracted by the degeneracies in this range.

We have explored using a larger set of input spectrum, by including
reddenned versions of all our templates, to test whether a more
complete space of SEDs would cause our fit to choose a more realistic
subspace. In addition, we have tried removing the increased resolution
of the template SEDs around the 4000 \AA break. Neither of these tests
yielded better results.

Our view is that the ideal solution to this problem would be to
minimize $\chi^2$ under the constraint that none of the original
Bruzual-Charlot spectra defining our SED space contribute negatively
to the SED fit. Such a solution would almost certainly have better
properties than our linear approach here: it would almost certainly
avoid the degeneracies, it would yield reasonable fit SED over a
larger wavelength range, it would associate a reasonable
star-formation history with each object (which could be used to
evolution-correct the magnitudes).

However, we have not tackled this approach yet. We have implemented
(and used for the figures shown here) a temporary fix for those
interested in galaxies at $z=0.3$, which is to artificially constrains
the fit as follows. We pretend that there is an extra band between $g$
and $r$, which has a FWHM of 2000 \AA\ and a center of 5500 \AA. The
``measured'' value we use in this band is the mean of the $g$ and $r$
fluxes, and is given $4\sigma$ significance. This extra constraint
changes the K-corrections for galaxies outside the range $0.27 < z <
0.33$ by far less than 1 percent. The constraint helps to clean up the
fits but it is not sufficient; the coefficients are still clearly
wrong around $z=0.3$. It helps considerably to $K$-correct to $z=0.3$,
which we recommend for the moment for studies of LRGs; this minimizes
the corrections which are performed using our most unphysical fits.

Remember, you really only need to worry about this effect in the SDSS
when you are dealing with the LRGs, for galaxies in the range about
0.27 < z < 0.33. The Main Sample results are not significantly
affected at all.  Furthermore, one can still observe a galaxy at
$z=0.1$ and reliably infer what it would look like at $z=0.3$; it is
only the reverse process which is difficult.

%It is clear from Figure \ref{lrg_colors_plot} that $z=0.3$ is a
%special redshift regime where our fits become unphysical. In fact, it
%is easy to demonstrate that the diagonal spur (going from the upper
%left to the lower right) around the red dot in the upper right panel
%of Figure \ref{k_coeffdist_plot} is due almost completely to galaxies
%at $z\sim 0.3$. The spur arises because the 4000 \AA\ break lies in
%the gap between $g$ and $r$ at $z=0.3$, visible in Figure
%\ref{response_sdss}. Our algorithm has strong gradients in the
%spectrum around the 4000\AA\ break, which is necessary to properly
%model galaxy spectra. In the presence of these sharp gradients, the
%gap in the bands causes certain combinations of the template spectra
%to be degenerate at around $z=0.3$.

%The question arises of what to do about this problem. One solution is
%to simply work around it. It is clear we cannot use the fits presented
%here to $K$-correct galaxies which are at $z=0.3$ to other redshifts
%reliably --- their coefficients $a_k$ are unusual and probably
%unphysical. However, nothing prevents one from $K$-correcting galaxies
%at other redshifts {\it to} $z=0.3$. If you are considering the LRG
%sample, the median redshift is close to $z=0.3$ anyway, so this is a
%viable solution (and the one adopted in Figure \ref{lrg_colors_plot}).

%A second solution is to place extra constraints on the fit which
%prevent the spur in the coefficients. One can imagine fitting the
%distribution in the $(a_1/a_0)$-$(a_3/a_0)$ plane with a simple model,
%such as a multigaussian fit, and using the resulting fit as a prior in
%one's fit (with gaussians, the extra constraints would be linear and
%thus easily accounted for). One could design these constraints to be
%strong enough (at least near the redshifts in question) to prevent the
%spur. However, the disadvantage of this solution is that for some
%scientific projects the change in the galaxy SED is exactly what one
%is trying to measure. In these cases, it may be inappropriate to be
%putting explicit {\it a priori} constraints on exactly what one is
%trying to measure.

%A third solution, not adopted here, is to place stronger constraints
%in the SED fit on the form of the 4000 \AA\ break. One obvious way of
%doing this is to use the spectra to constrain the size of the break
%(making it a weak constraint to prevent biases). This constraint could
%be applied in a linear fashion to retain the speed of our method ---
%except for the fact that you would have to read in the spectra to
%apply the $K$-correction. For simplicity, and because the aforemention
%work-around is so easy to obey, we have not implemented this solution
%here.
